Learning Auditory Similarities Between Instrumental Playing Techniques

### 2020-05-22 paper/paper.tex

The European symphonic orchestra encompasses four families of instruments: strings, woodwinds, brass, and percussion.
In this article, we focus on the former three, and leave the question of learning auditory similarities between percussion instruments to future research.
The reason behind this choice is that most percussive instruments do not allow much variation in terms of pitch or intensity, if at all; thus reducing the amount of intra-technique variability in the time--frequency domain.
Furthermore, although percussive instruments certainly exhibit a rich diversity of playing techniques (we refer to \cite{peinkofer1976book} for a survey), these are rarely shared with pitched instruments.
In comparison, techniques for sustained tones such as \emph{vibrato} or \emph{tremolo} permeate throughout a vast instrumentarium, \eg{} bowed strings and reed alike.
We refer to \cite{wu2018taslp} and \cite{pearce2019appliedsciences} for reviews of the recent literature on the timbre modeling of percussive instruments, from the standpoints of MIR and music cognition respectively.

@@

\subsection*{Expert pre-screening}
Secondly, we reduce the number of IMTs in our study by focusing on those which are deemed to be most relevant.
Here, we define the relevance of an IMT as the possibility of imitating it by means of another IMT from a different instrument.
One example of such imitation is the acoustic similarity between slap tonguing in reed instruments and a snap pizzicato in strings instruments.
To collect perceptual ratings of relevance, we recruited two professors in music composition at the Paris Conservatory (CNSMDP), and explicitly gave them our definition of relevance.
Each of them inspected the entire corpus of $235$ IMTs and annotated them in terms of relevance according to a Likert scale with seven ticks.
In this Likert scale, the value 1 (least relevant) denotes that the IMT under consideration has a timbre that is idiosyncratic of the instrument, and that therefore, it is unlikely that humans will pair it with IMTs from other instruments.
Conversely, the value 7 (most relevant) denotes that the IMT under consideration has a strong similarity with IMTs from one or several other instruments.

Once both experts completed their annotations, we retained all IMTs whose average relevance score was judged equal to 3 or higher.
At the end of this stage of pre-screening, the list of IMT reduces from 235 to $N=78$.
Tables 1 and 2, in the Appendix section of this paper, give a full list of these 78 highly relevant IMTs.
It is worth noting that, according to both experts, the timbre of the accordion was judged too idiosyncratic, regardless of playing technique, to be relevant for this experiment.
Indeed, the accordion is the only instrument in the aforementioned list of 
instrument to have free reeds, keyboard-based actuation, or handheld airflow.
Consequently, regardless of mute and technique, the set of instruments $\mathcal{I}$ in our study contains 15 elements.

>>>>

The European symphonic orchestra encompasses four families of instruments: strings, woodwinds, brass, and percussion.
In this article, we focus on the former three, and leave the question of learning auditory similarities between percussion instruments to future research.
The reason behind this choice is that most percussive instruments do not allow much variation in terms of pitch or intensity, if at all; thus reducing the amount of intra-technique variability in the time--frequency domain. \ml{preciser le contexte, un joueur de tabla ne serait pas d'accord, j'enleverai ces 2  phrases, ... and refer to ...}
Furthermore, although percussive instruments certainly exhibit a rich diversity of playing techniques (we refer to \cite{peinkofer1976book} for a survey) these are rarely shared with pitched instruments.
In comparison, techniques for sustained tones such as \emph{vibrato} or \emph{tremolo} permeate throughout a vast instrumentarium, \eg{} bowed strings and reed alike.
We refer to \cite{wu2018taslp} and \cite{pearce2019appliedsciences} for reviews of the recent literature on the timbre modeling of percussive instruments, from the standpoints of MIR and music cognition respectively.

@@

\subsection*{Expert pre-screening}
Secondly, we reduce the number of IMTs in our study by focusing on those which are deemed to be most relevant.
Here, we define the relevance of an IMT as the possibility of imitating it by means of another IMT from a different instrument.
One example of such imitation is the acoustic similarity between slap tonguing in reed instruments and a snap pizzicato in strings instruments.
To collect perceptual ratings of relevance, we recruited two professors in music composition at the Paris Conservatory (CNSMDP), and explicitly gave them our definition of relevance.
Each of them inspected the entire corpus of $235$ IMTs and annotated them in terms of relevance according to a Likert scale with seven ticks.
In this Likert scale, the value 1 (least relevant) denotes that the IMT under consideration has a timbre that is idiosyncratic of the instrument, and that therefore, it is unlikely that humans will pair it with IMTs from other instruments.
Conversely, the value 7 (most relevant) denotes that the IMT under consideration has a strong similarity with IMTs from one or several other instruments \ml{available in the database}.

Once both experts completed their annotation, we retained all IMTs whose \ml{relevance score averaged over experts} was judged equal to 3 or higher.
At the end of this stage of pre-screening, the list of IMT reduces from 235 to $N=78$.
Tables 1 and 2, in the Appendix section of this paper, give a full list of these 78 highly relevant IMTs.
It is worth noting that, according to both experts, the timbre of the accordion was judged too idiosyncratic, regardless of playing technique, to be relevant for this experiment.
Indeed, the accordion is the only instrument in the aforementioned list of
instrument to have free reeds, keyboard-based actuation, or handheld airflow.
Consequently, regardless of mute and technique, the set of instruments $\mathcal{I}$ in our study contains 15 elements.
